%% -*- coding:utf-8 -*-
\documentclass[ number=??
                ,series=lnls,
                ,isbn=xxx-x-xxxxxx-xx-x,
                ,url=http://langsci-press.org/catalog/book/14,
	        ,output=long    % long|short|inprep              
	        %,blackandwhite
	        %,smallfont
	        ,draftmode  
		  ]{langsci}                          

          
\hypersetup{pdfdisplaydoctitle=true}

\usepackage{xspace}
\newcommand{\latex}{\LaTeX\xspace}



% just for the XeLaTeX logo
\usepackage{dtklogos}\newcommand{\xelatex}{\XeLaTeX\xspace}
\newcommand{\bibtex}{\BibTeX\xspace}

\usepackage{lsp-makros}

\usepackage{german}\selectlanguage{USenglish}

\usepackage{lsp-eng-hyp}

\usepackage{abbrev}


% AVMs
\usepackage{lsp-styles/avm}
\avmfont{\sc} 
\avmvalfont{\it}
      
% OT pointing hand
\usepackage{pifont}
\newcommand{\hand}{\ding{43}}

% OT tableaux                                                
\usepackage{pstricks,colortab}   

% DRS package by Alexis Dimitriadis
%\usepackage{drs}

\usepackage{tabularx}


\usepackage{tikz-qtree}
% has strange side effects
%\tikzset{every tree node/.style={align=left, anchor=north}}
\tikzset{every roof node/.append style={inner sep=0.1pt,text height=2ex,text depth=0.3ex}}

\usepackage{jambox}
 
\usepackage{lipsum}  

\setlength{\marginparwidth}{1.5cm}
\usepackage[textsize=tiny,textwidth=1.5cm]{todonotes}
\newcommand{\todostefan}[1]{\todo[color=green!40]{#1\xspace}}
\newcommand{\inlinetodostefan}[1]{\todo[color=green!40,inline]{{\normalsize #1}}}


% Chinese
\usepackage[indentfirst=false]{xeCJK}
\setCJKmainfont{SimSun}

% bidirectional text and support for Arabic/Persian
%% \usepackage{fontspec}
\newfontfamily\Parsifont[Script=Arabic]{XB Niloofar}
%\usepackage{bidi}
\usepackage{lsp-bidi}
\newcommand{\PRL}[1]{\RL{\Parsifont #1}}
%\TeXXeTOff
\usepackage{lsp-gb4e}



\title{How to Write a Book for Language Science Press}  
\subtitle{Guidelines for Authors and \latex Recommendations}
\BackTitle{How to Write a Book for Language Science Press}
\BackBody{This book contains the guidelines for Language Science Press authors. For those who want
  to help keeping the production costs low and therefore decided to use \latex it also contains
  descriptions of packages that can be used for typesetting trees, Attribute Value Matrices, OT-tableaux, Categorial
  Grammar proofs, LFG analyses, and much more. The setup of typesetting script with special fonts as
  for instance right to left scripts like Arabic is explained. The \latex chapter also contains sections
  concerning the efficient workflow in professional typesetting environments using \latex.

Stefan Müller is an experienced \latex user who has typeset four published books and several book
manuscripts and journal articles.
}

\dedication{This book is dedicated to everybody who cannot afford to buy books by profit oriented publishers.}

\author{Stefan Müller}



\makeatletter
\def\verbatim@font{\scriptsize\ttfamily}
\makeatother
         
%\renewcommand{\eachwordone}{\it}
%\renewcommand{\exfont}{\it}
%\def\exfont{\it}

\begin{document}               
         
                                                                           
                                  
\maketitle                

%\frontmatter

\chapter*{Preface}

%\lipsum[3-10]  

The purpose of this book is twofold: it contains a guideline with some style recommodations for all
authors. The second part is for authors who use \latex or who want to learn \latex in order to
support \lsp. The second part is also a reference for those who volunteered to help typesetting
manuscripts that were not submitted in \latex. See \citew{MuellerOA} and \citew{MH2013a} for an overview of the
general setup of the project.

\section*{Acknowledgements}



%% David Reitter danke ich für die \LaTeX"=Makros für Combinatorial
%% Categorial Grammar, Mary Dalrymple und Jonas Kuhn für die LFG"=Makros und Beispielstrukturen, und
%% Laura Kallmeyer für die \LaTeX"=Quelltexte der meisten TAG"=Analysen. Ohne diese
%% Originalmakros/-texte wären die Darstellungen in den jeweiligen Kapiteln sicher nicht halb so schön
%% geworden. 

This book is typeset with \xelatex. We thank the \latex developers for their work and the members of the German \textit{German
  Language TeX Users Group Communication List} and those replying at \url{http://tex.stackexchange.com} for many usefull hints and suggestions.

I thank Matthias Hüning\aimention{Matthias H{\"u}ning} for comments on an earlier version of this document and Corinna Handschuh\aimention{Corinna Handschuh}
and Francesco Cangemi\aimention{Francesco Cangemi} for being the first ones using the new \latex classes and providing feedback
to us.

\bigskip

\noindent
Berlin, \today\hfill Stefan Müller


\tableofcontents      

\mainmatter         

\include{lsp-setup}
\include{lsp-authors}
\include{lsp-latex}
\include{lsp-publication}


\backmatter

\bibliography{lsp-abbrev,lsp-guidelines}

%\cleardoublepage

\clearpage
\pdfbookmark[0]{Index}{Index}
\pdfbookmark[1]{Index of Names}{Index of Names}
\printindex[aut]
%\addcontentsline{toc}{chapter}{Verzeichnis der Sprachen}
\pdfbookmark[1]{Index of Languages}{Index of Languages}
\printindex[lan]
%\addcontentsline{toc}{chapter}{Sachregister}
\pdfbookmark[1]{Index of Subjects}{Index of Subjects}
\printindex

                              
\end{document}
      
%      <!-- Local IspellDict: en_US-w_accents -->
