\documentclass{article}
\usepackage{libertine,hyperref,tikz}
 \usepackage[
papersize={170mm,240mm}
,top=27.4mm
,inner=20.5mm,
,outer=24.5mm
%,showframe,pass
,marginparwidth=50pt
]{geometry}
\begin{document}
	\section*{This is a draft}
	Dear reader,\\\\
	you have received a draft from a book or chapter from Language Science Press.
	
	\begin{itemize}
		\item At Language Science Press, we have specific markup requirements for the usage of quotation marks and italics. Some changes may have already been made to your text in case the usage of quotation marks and italics did not comply with our standards. Please do check the \href{https://langsci.github.io/gsr/GenericStyleRulesLangsci.pdf}{Generic Style Rules} and our \href{http://langsci.github.io/guidelines/latexguidelines/LangSci-guidelines.pdf}{house rules} for these rules.	
		\item Black rectagles \tikz \fill (0,0) rectangle (.15,.25); in the page margin indicate that the content did not fit the page. For figures and tables, this is often due to them being too large. If black rectangles appear in the running text, this is mostly due to the line algorithm being unable to find a satisfactory hyphenation solution. This may be due to technical or foreign phrases in the respective line. In this case, please provide the correct hyphenation in \verb|localhyphenation.tex|. As \LaTeX does not break phrases that contain a hyphen (like \textit{Alor-Pantar}), we have to insert possible breaks manually (as in \verb+Alor-Pan\-tar+). Your copy editor will be happy to insert these breaks, you'll only need to send a list.
		\item Notes in orange are comments that the editor needs input for. In most cases, these notes will be addressed to the author of the book or chapter or the editor of the volume. If you know the fix for the problems indicated by the notes, please write to your us.
		\item If a cited source appears in bold and without spaces (e.g.\ \textbf{Chomsky2001}), this bibliographical entry could not be found in the \verb|.bib| file.
		\item A bold-faced \textbf{??} indicates that the target for a reference (to a table, figure, example, page, etc.) could not be found. Please check the respective \verb|\label{}| and \verb|\ref{}| commands in your text.
	\end{itemize}
\end{document}